---
title: "HOWTO generate biocViews HTML"
author: "S. Falcon and V.J. Carey"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteIndexEntry{Quick-start Guide}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
output:
  BiocStyle::html_document:
    number_sections: no
    toc: yes
    toc_depth: 2
editor_options: 
  markdown: 
    wrap: 72
---


```{r include=FALSE}

if(!require(biocViews)){
    install.packages("biocViews")
    library(Biobase)
}

```

# Overview

The purpose of *biocViews* is create HTML pages that categorize packages
in a Bioconductor package repository according to terms, or *views*, in
a controlled vocabulary. The fundamental resource is the VIEWS file
placed at the root of a repository. This file contains the complete
DESCRIPTION file contents for each package along with additional meta
data describing the location of package artifacts such as archive files
for different platforms and vignettes.

The standard behavior of the view generation program is to query the
repository over the internet. This package includes a static sample
VIEWS file so that the examples in this document can run without
internet access.

# Establishing a vocabulary of terms

We use `dot` to describe the vocabulary. For details on the `dot`
syntax, see <http://www.graphviz.org/doc/info/lang.html>.

```{r Establishing a vocabulary of terms, echo=TRUE, message=FALSE, warning=FALSE}

vocabFile <- system.file("dot/biocViewsVocab.dot", package="biocViews")
cat(readLines(vocabFile)[1:20], sep="\n")
cat("...\n")

```

The dot description is transformed to a GXL document using `dot2gxl`, a
tool included in the graphviz distribution. The GXL is then converted to
a *graphNEL* instance using `fromGXL` from the *graph* package. There is
a helper script in the root of the *biocViews* package called
`updateVocab.sh` that automates the update process if the required tools
are available. The script will also attempt to dump the ontology graph
into a local SQLite database using tools from *DBI* and *RSQLite*. The
information in this database can be used to create a dynamic HTML
representation of the graph by means of a PHP script.

The definition of the vocabulary lacks a notion of order. Since the
purpose of the vocabulary is primarily for display, a valuable
improvement would be to use graph attributes to allow the ordering of
the terms.

Another missing piece is a place to put a text description of each term.
This could also be achieved using graph attributes.

## Use Case: adding a term to the vocabulary

To add a new term to the vocabulary:

1. edit the *dot* file `dot/biocViewsVocab.dot` and add the desired term. Note that terms cannot contain spaces and that the underscore character, `_`, should be used instead.
2. ensure that R and dot2gxl are on your PATH.
3. cd into the biocViews working copy directory.
4. run the updateVocab.sh script.
5. reinstall the package and test that the new term is part of the vocabulary. In short, you will load the data using `data(biocViewsVocab)` and check that the term is a node of the graph instance.
6. commit changes to svn.

## Use Case: updating BioConductor website

This is for BioConductor web administrator:

1. update local copy of biocViews using `svn update`.
2. find the correct instance R that is used to generate HTML pages on BioConductor website, and install the updated `biocViews`.
3. re-generate the related HTML packages by using `/home/biocadmin/bin/prepareRepos-*.sh` and `/home/biocadmin/bin/pushRepos-*.sh`.

# Querying a repository

To generate a list of *BiocViews* objects that can be used to generate
HTML views, you will need the repository URL and a graph representation
of the vocabulary.

There are three main Bioconductor package repositories: a software
repository containing analytic packages, an annotation data repository,
and an experiment data repository. The vocabulary of terms has a single
top-level node, all other nodes have at least one parent. The top-level
node, *BiocViews*, has three children that correspond to the three main
Bioconductor repositories: *Software*, *AnnotationData*, and
*ExperimentData*. Views for each repository are created separately using
`getBiocSubViews`. Below, we demonstrate how to build the *Software* set
of views.


```{r Querying a repository, echo=TRUE, message=FALSE, warning=FALSE}

data(biocViewsVocab)
reposPath <- system.file("doc", package="biocViews")
reposUrl <- paste("file://", reposPath, sep="") 
biocViews <- getBiocSubViews(reposUrl, biocViewsVocab, topTerm="Software")
print(biocViews[1:2])

```



To query the currently available vocabulary terms, use function
`getSubTerms` on the *graphNEL* object `biocViewsVocab`. The second
argument of this function takes a character of the base term for which
all subterms should be returned. For a complete list use
`term="BiocViews"`.

```{r getSubTerms, message=FALSE, warning=FALSE}

getSubTerms(biocViewsVocab, term="Technology")

```



# Generating HTML

By default, the set of HTML views will link to package description pages
located in the html subdirectory of the remote repository.

```{r Generating HTML, echo=TRUE, message=FALSE, warning=FALSE}

viewsDir <- file.path(tempdir(), "biocViews")
dir.create(viewsDir)
writeBiocViews(biocViews, dir=viewsDir)
dir(viewsDir)[1:2]
 
```

